---
title: "qualtrics_spss"
editor: source
---

Using the tidyverse package and the haven package. {haven} is used to import `.sav` files

```{r}
library(tidyverse)
library(haven)
library(tidytext)
```

## Import data

The following is test data randomly generated at Qualtrics. From Qualtrics, the same data were exported: once as a CSV file; another copy as a .sav (label \[SPSS \| SAS \| Stata\] ) file. When the Qualtrics data are downloaded as CSV, survey developer can download data as numeric response or categorical-text (option: "use choice text".)

```{r}
my_student_survey_spss_sas <- read_sav("data/student+satisfaction+-+test+data_April+11,+2023_17.30.sav") |> 
  janitor::clean_names()
```

## Wrangle labeled (SPSS SAS) data

Create one labeled column to a character column of categories

```{r}
my_student_survey_spss_sas |> 
  mutate(Q2_label = as_factor(q2), .after = q2)
```

Convert all labeled columns to categorical columns

```{r}
my_student_survey_spss_sas <- my_student_survey_spss_sas |> 
  mutate(across(is.labelled, ~ as_factor(.x), .names = "{.col}_label"))
```

We can also pivot the data to a long format. This will make it easier to iterate over mulitple visualizations

```{r}
my_df_tall <- my_student_survey_spss_sas |> 
  select(response_id, contains("label")) |>
  pivot_longer(-c(response_id, q15_label), names_to = "question", values_to = "response")
my_df_tall
```

## Example Viz

By importing data as labeled categories, it's easier to use that information in a visualization such as the following

### data tables

First, let's look at the data frames in the wide format

```{r}
my_student_survey_spss_sas
```

And now let's make a faceted bar chart from the tall formatted data

```{r}
#| fig-width: 12
#| fig-asp: 1.2

# NOTE:  Render this code using the render button.  It looks better in the rendered document


my_df_tall |> 
  filter(question != "status_label",
         question != "finished_label") |> 
  ggplot(aes(y = response)) +
  geom_bar() +
  facet_wrap(vars(question), scales = "free", ncol = 3)
  
```



```{r}
#| fig-width: 12
#| fig-asp: 1.2
#| echo: false


# FOR README ##


my_df_tall |> 
  filter(question != "status_label",
         question != "finished_label") |> 
  ggplot(aes(y = response)) +
  geom_bar() +
  facet_wrap(vars(question), scales = "free", ncol = 2) +
  labs(x = NULL, y = NULL, title = "Labeled categories are sorted in imported factor order") +
  theme(plot.title.position = "plot")

ggsave("output/viz_as_imported_labeled_data.svg", width = 12, height = 14.5)
```

